### 2009-08-02 reviewers/review-response.tex
basics of review-response in place

% -*- LaTeX -*-
\documentclass[letterpaper,12pt]{article}

\newenvironment{response}
{\begin{quotation} \bf}
  {\end{quotation}}

\begin{document}

\section{Comments for the Author: Reviewer \#1}
\label{sec:reviewer1}

This submission on Dependency Pair Match MT metrics is technically and
structurally sound, thorough, and well-written.  Some specific
comments/questions below, in order of appearance in the paper.
\begin{response}
  Thank you for this feedback. Our responses are below.
\end{response}
As per the submission guidelines, submissions must be limited to 12
pages of content, with an additional page for references.  The current
version has 16.5 pages plus one page of references, so significant
shortening will be necessary.  I find the overall balance of sections
adequate, so I would recommend shortening across the board rather than
in one particular section.
\begin{response}
  We have done revisions to compress and simplify the writing
  throughout the paper to address the length concerns.
\end{response}
While likely known by most of the community, I would recommend
expanding all acronyms the first time (LFG, PCFG, ...).

In section 2, ``bag of dependent word'' should be ``bag of dependent
word*s*''.
\begin{response}
  We believe that we have emended these.
\end{response}
Can you expand on the reasoning behind the modifications to the
Charniak parser rules described in section 3.1?
\begin{response}
  We hope that our examples (and some additional explanatory sentences
  in section 3) have addressed this concern.
\end{response}
In section 4.1, while the corpus is referenced and the information can
be looked up, it would still be helpful to include directly in the
paper a brief description of the adequacy and fluency measures.  In
particular, what kind (granularity) of scale was used for the
judgments needs to be known in order to be able to assess the reported
correlations.  A brief description of the data genre(s) would be
useful to have at hand as well.  Also, I assume you correlate at the
*segment* level, but this is not explicitly stated.
\begin{response}
  Section 4 has now included several additional sentences to clarify
  these points.
\end{response}
The correlations reported in section 4 (tables I-III) are overall
quite low (also compared to EDPM's r correlations for Chinese source
data in NIST's MetricsMATR challenge).  Could you comment on this?
\begin{response} [TO DO]
\end{response}
For ease of lookup, please include in the table headings for tables
I-III that the correlations are with human judgments of adequacy and
fluency

In section 5.1, ``EDPM's strengths, relative to the other measures, are
particularly clear in the unstructured domains (wb and bc).''  This
statement does not seem to hold for EDPM vs. BLEU for the wb genre in
table Va; the correlation reported is the same for the two metrics?

\section{Comments for the Author: Reviewer \#2}
\label{sec:reviewer2}

This article presents a method for evaluating MT system hypotheses
that is based on complete and partial dependency triple matches
between hypothesis and reference translation.

The paper is well motivated, clearly written, and presents interesting
results. On my opinion, it is suitable for publication with minor
revisions. My main points of criticism are:

- It is considerably too long. Papers are supposed to be 12 pages +
 references; this paper has more than 16 pages. I recommend
 (a) shortening the introduction; (b) shortening the discussion
 of the "model space" in Sections 2/3 somewhat; (c) removing the
 third experiment (Section 6), whose contribution is unclear to
 me, altogether. (see below for details)

- On the other hand, the paper does not motivate at all what the
 intuition behind the use of dependency triple match is. Why
 should we assume in the first place that dependency triples
 predict human judgments better than (e.g.) n-grams?

- Finally, the paper does not give a single example. In a paper
 about a novel evaluation metric, I would like to get a feeling
 for what it is that the new metric does better than existing
 metrics. (This can be addressed together with the previous point.)

Detail comments:

- The discussion on page 2 reads nicely, but is probably too
 extensive -- in particular for a paper aimed at a special
 issue where there will presumably be an overview article
 that discusses relevant previous work.

- p.2, publically -> publicly?

- p.3, "In section 2 etc." I'm not sure that a 12-page paper requires
 a table of contents.

- p.3, reference to Owczarzak et al.'s work. In particular since their
 work has also appeared in the MT journal (please update the
 reference, btw!), it would be good to expand on the similarities and
 differences of your work and theirs.

-p.4 I'm not so sure that I believe in the benefit of overweighting.
 In standard dependency graphs, for example, modifications are usually
 dependendents of the head and are thus "underweighted" by your scheme,
 but modifiers like negations etc. can considerably change the meaning.
 (You write "we argue" but I don't see an argument, just a claim --
  details of counts for an example might be helpful.)

-p.5 "When multiple decompositions etc." As far as I can see, you
 get consistently better scores for F[] than for $\mu_{PR}[]$ throughout
 your experiments.  I'd suggest reducing the discussion of $\mu$ here
 and in Section 3.2 with a short explanation that $\mu$ has been found to
 work worse than F.

- p.6 "LFG...heavily knowledge-engineered parser". I think you're being
 overly harsh on the LFG-based work here. As the title of the paper you
 cite says, you're talking about an "automatically acquired PCFG-based
 LFG approximation" that's learned from a treebank, so you are really
 very much in the same boat.

- p.7 "Score combination" - see comment above
     "Tuned, like the ..." - ungrammatical sentence?

- Evaluation (p. 8) In principle I like the idea of teasing apart
 sentence difficulty and translation quality. However, I have two
 reservations: (1), you are losing comparability of scores across
 sentences; and (2), you are committing to a scenario where you
 always have access to a sufficient number of translations in order
 to estimate the mean score. In other words, in the limit of one
 system, you do not predict anything.  These issues should be
 mentioned and discussed.  (Note about presentation: you don't
 introduce the variable t in Equation 3 -- I guess this is the
 segment?)

 More generally, I would like to see a motivation for the evaluation
 you perform. You seem to largely follow Owczarzak et al's
 evaluation, e.g. in the choice of data (Section 4.1), but you depart
 from it in some crucial respects (e.g. the definition of the
 dependent variable by mean subtraction). Can you provide a clearer
 description of why each experiment is set up in the way it is?

- p.10 Am I right in reading Table II to say that F always outperforms
 BLEU, but $\mu$ never does? Maybe it would be good to re-mention the
 BLEU results in Table II.

- p.12 Can you elaborate on why you want to weight per-sentence
 correlations by sentence length? I don't feel that this
 appropriate. After all, I'd think that you are not \emph{more} interested
 in longer sentences, but you want each sentence to have the same
 contribution to the analysis. I assume the problem is that there is
 considerably more variance for shorter sentence, which is alleviated
 by length-based weighting, but to me this reads like a cover-up. If
 you have good arguments for wanting long sentences to count more in
 the analysis, I'd be interested to hear them.

- Section 6. To my mind, this is the least convincing section of the
 paper. The experiments in Section 4 have already shown that a simple
 F-score-based combination of the different matches does better than
 considering individual components. While I agree that this might be
 a weighting problem, this experiment shows basically a null result.
 From what I can gather from Figure 4, TERp on its own does
 (significantly?) better than EDPM, and their combination doesn't
 really improve a lot. I acknowledge the positive result that there
 is no \emph{decrease} in performance either, but then you're using quite
 a lot of data (50\% = 2500 sentences) for tuning so that is to be
 expected. I would very much like to see how robust the tuning
 process is to smaller development sets, but that would require even
 more space.

 In sum, I'd suggest replacing Section 6 with some actual examples
 and more discussion of the methodical points I have suggested.
\end{document}